---
title: "dlookr demo"
output: html_notebook
---

Install dlookr in r:

```{r}
#install.packages("dlookr")
# If you haven't already installed dplyr, do that now, as we will use dplyr in conjunction with 
# install.packages("dplyr")
```

Bring dlookr and dplyr into memory by calling the libraries:

```{r}
library(dplyr)
library(dlookr)
```

Load in your data:

(We'll be using the Ames training data from Kaggle for this example.)

```{r}
load("ames_train.Rdata")
```

Let's take a look at our dataset:

```{r}
dim(ames_train)
names(ames_train)
head(ames_train)
```

The codebook for the data is saved as "ames_data_codebook.txt"

***
#Data Quality Diagnosis

Now let's dive into dlookr to diagnose our data quality:

* diagnose() allows you to diagnosis a variables in a data frame. Like function of dplyr, the first argument is the tibble (or data frame). The second and subsequent arguments refer to variables within that data frame.
* diagnose_category() provides detailed diagnostic information for categorical variables.
* diagnose_numeric() provides detailed diagnostic information for numeric variables.
* diagnose_outlier() and plot_outlier() provide information and visualization of outliers.

The variables of the tbl_df object returned by diagnose () are as follows.

* variables : variable name
* types : the data type of the variable
* missing_count : number of missing values
* missing_percent : percentage of missing values
* unique_count : number of unique values
* unique_rate : rate of unique value. unique_count / number of observation

```{r}
diagnose(ames_train)
```

By using dplyr, variables including missing values can be sorted by the weight of missing values.:

```{r}
ames_train %>%
  diagnose() %>%
  select(-unique_count, -unique_rate) %>% 
  filter(missing_count > 0) %>% 
  arrange(desc(missing_count))
```

The missing data makes sense. Homes without a pool wouldn't have an entry for Pool.QC, likewise with Misc.Feature, Alley, Fence, Fireplace.Qu, etc. Refer to the codebook to understand each variable better.

You can also diagnose the numeric values:

```{r}
diagnose_numeric(ames_train)
```

Here, for example, we see PID is not a predictor or meaningful value:

PID (Nominal): Parcel identification number  - can be used with city web site for parcel review. 

We can also see how many outliers there are in each variable.

Likewise, we can also look at the categorical variables by using the diagnose_category command:

```{r}
diagnose_category(ames_train)

```

And you can look at only the missing values by:

```{r}
diagnose_category(ames_train) %>% 
  filter(is.na(levels))
```

And you can review the data outliers:

```{r}
diagnose_outlier(ames_train) %>% 
  filter(outliers_cnt > 0) 
```

You can also easily plot the outliers using the plot_outlier() command:

```{r}
ames_train %>%
  plot_outlier(area) 
```

Use the function of the dplyr package and plot_outlier() and diagnose_outlier() to visualize anomaly values of all numeric variables with an outlier ratio of 0.5% or more.:

```{r}
ames_train %>%
  plot_outlier(diagnose_outlier(ames_train) %>% 
                 filter(outliers_ratio >= 0.5) %>% 
                 select(variables) %>% 
                 unlist())
```

Or you can skip all that and get dlookr to output a complete diagnostic report:

```{r}
ames_train %>%
  diagnose_report(output_format = "html", output_file = "Diagn.html")
```


***
#Exploratory Data Analysis

Once you've diagnosed the data quality, the next step is to explore data to understand the data and find scenarios for performing the analysis.

The following is a list of the EDA functions included in the dlookr package.:

* describe() provides descriptive statistics for numerical data.
* normality() and plot_normality() perform normalization and visualization of numerical data.
* correlate() and plot_correlate() calculate the correlation coefficient between two numerical data and provide visualization.
* target_by() defines the target variable and relate() describes the relationship with the variables of interest corresponding to the target variable.
* plot.relate() visualizes the relationship to the variable of interest corresponding to the destination variable.
* eda_report() performs an exploratory data analysis and reports the results.

Describe gives summary statistics for each variable. Using dplyr, you can sort and detail them as needed. For example, you can sort by skewness or any other factor:

```{r}
ames_train %>%
  describe() %>%
  select(variable, skewness, mean, p25, p50, p75) %>% 
  filter(!is.na(skewness)) %>% 
  arrange(desc(abs(skewness)))
```

You can report on all variables or select variables and use the group_by() function ...

You can also check the normality of the data with the dlookr normality() test:

```{r}
ames_train %>%
  normality() %>%
  filter(p_value <= 0.01) %>% 
  arrange(abs(p_value))
```

Here we run normality on the dataset, show only the results with p-value less than or equal to 0.01 and sort by absolute p-value.

And you can plot the distributions for a visual reference. (You can plot all or select variables.)

```{r}
plot_normality(ames_train, area, Full.Bath)
```

Notoice the normality plots automatically show the distribution of the log and sqrt function.

***
## Bivariate data EDA
Calculation of correlation coefficient using correlate()
Correlate() finds the correlation coefficient of all combinations of ames_data numerical variables as follows:

```{r}
correlate(ames_train)
```

Then you can plot the correlations with a simple function. (I've only selected a couple of variables here to make the graph easier to read.)

```{r}
plot_correlate(ames_train, area, price)
```

***
##EDA Using Target Variable

You can also perform EDA based on a target variable, which requires creating a target_by class object. target_by() creates a target_by class with an object inheriting data.frame. 

target_by() is similar to group_by() in dplyr which creates grouped_df. The difference is that you specify only one variable.

*This is a bit out of the scope of this demo, but documentation is available on the cran page for dlookr.

***
Finally, you can also run a single EDA report just as with the data quality diagnosis report. 

```{r}
ames_train %>%
  eda_report(target = price)
```

***

The final tool in dlookr is data transformation. That, too, is a topic for another day.

https://cran.r-project.org/web/packages/dlookr/vignettes/transformation.html




